Toward GPU-accelerated Traffic Simulation and Its Real-Time Challenge

### 2012-11-04 draft.tex
futher minor fix
updated for the camera-ready

\begin{abstract}
 Traffic simulation is a growing domain of computational physics.
 There are many applications in our life and industry that benefit
 from traffic simulation.
 A core challenge of this science research, however, is posed due to
 an unbounded scale of computation.
 In this paper, we explore an advantage of using the graphics processing
 unit (GPU) to support its computation.
 Two schemes are presented to study how to maximize GPU
 performance in the context of traffic simulation.
 Our experimental results show that even the most basic GPU
 implementation based on our schemes accelerates simulation by about
 five times over the original CPU implementation.
 We also demonstrate that additional orders-of-magnitude
 improvements could be achieved by overcoming the current hardware
 limitation.
\end{abstract}

@@

Transportation systems underlie our life and industry, as part of
societal infrastructure.
Traffic flow is particularly a major factor that influences economy.
According to the Japanese government report, economic losses due to
traffic congestion could reach one hundred billion dollars per year in
Japan, through degradation of transport efficiency, energy consumption,
and environmental poisoning.
Albeit a major issue of transportation systems, the mechanism of traffic
congestion is not well-explained in the literature.
What is particularly important is ``phantom jam'', which often happens
on freeways without any accident.
This phantom jam should be avoided by science, given that it is
a physical phenomenon but not one caused by human errors.
However, we have not been aware of scientific and engineering solutions
to such real-world problems yet.

@@

Applying a large number of agents to simulation using the above formula,
it is apparent that computational workload increases exponentially.
In fact, the above formula considers only one dimension, which restricts
applications of simulation to freeway traffic flows, powder flows,
molecular motors, and so on.
Making it multi-dimensional further increases computational workload,
while allowing more complicated simulations, such as traffic networks,
internet packet flows, evacuation routes, and herd formations of
animals, to use similar optimal velocity models.
Given a scale of million agents in the real world, traffic
simulation should be supported by powerful computer systems.

@@

This paper explores how to accelerate traffic simulation using the
state-of-the-art parallel computing technology.
In particular, we use the graphics processing unit (GPU), which
integrates hundreds of cores on a chip.
Recent GPUs are becoming more and more suitable for general-purpose
data-parallel applications.
The traffic simulation program used in this paper applies
Equation~\eqref{eqn:ov} to a large number of agents.
The resulting workload is highly data-parallel and compute-intensive,
which can be nicely offloaded on to the GPU.
We also identify the bottleneck in accelerating our traffic simulation
program using the GPU, and provide some insight into its solution.

The rest of this paper is organized as follows.
Section~\ref{sec:assumption} describes the basic assumption and
terminology used in this paper.
Section~\ref{sec:traffic_simulation} explains an overview of
traffic simulation and its literature.
Section~\ref{sec:gpu_implementations} presents our schemes of GPU
implementations for traffic simulation.
Section~\ref{sec:evaluation} evaluates the advantage of using our
schemes, and provides an insight into future work.
Section~\ref{sec:conclusion} concludes this paper.

@@

While we are intensively interested in CUDA and GPUs, the notion of
accelerating traffic simulation presented in this paper is widely
applicable to heterogeneous architectures, where CPUs and 
accelerators co-exist to complement each other. 
GPUs are currently the most well-recognized forms of accelerators, but
emerging alternatives include the Intel Many Integrated Core (MIC) and
AMD Fusion architectures.
However, their programming models are almost identical in that a master
thread running on the CPU manages the program flow, while parallelized
threads are offloaded on to accelerators to boost computing performance.
CUDA, OpenCL, HMPP, and AMP are all such programming languages that
support heterogeneous architectures under this concept.

@@

So far we have studied a single process of traffic flow.
In reality, however, there are multiple processes of traffic flow.
Real-time feedback from the real road conditions may also be
additionally required to deploy traffic simulation.
All these pieces of simulation must meet the quality-of-service (QoS)
requirement for users.
Imagine how many potential users exist in the real-world transportation
system.
We are facing a core challenge of next-generation real-time systems.

We have developed basic solutions to use the GPU in real-time
systems~\cite{Kato2011_3, Kato2011_2, Kato2011_1, Kato2012}.
Resource management with scheduling, synchronization, and memory
management of GPU applications has been addressed in our previous
work.
We believe that a grander vision of GPU-accelerated traffic
simulation will need an integration of cutting-edge real-time systems
technology and computational physics.

@@

In future work, we extend our scheme to multidimensional OV models.
This extension allows our research to be applied to a broader range of
problems beyond freeways traffic simulation.
We also plan to consider a specific approach to the integration of the
Imprecise Computation model into our scheme in order to facilitate
deployment of traffic simulation in the real world.
Another interesting direction in this line of work is to coordinate with
engineering-based simulation such as AutoMatrix~\cite{Mangharam2011}.

>>>>

\begin{abstract}
 Traffic simulation is a growing domain of computational physics.
 Many life and industrial applications would benefit from traffic
 simulation to establish reliable transportation systems.
 A core challenge of this science research, however, is its unbounded
 scale of computation.
 This paper explores an advantage of using the graphics processing
 unit (GPU) for this computational challenge.
 We study two schemes of maximizing GPU performance in the context of
 traffic simulation, and provide some basic experiments.
 The experimental results show that our GPU implementation improves
 simulation speed by five times over the traditional CPU
 implementation.
 We also discuss that additional orders-of-magnitude improvements could
 be achieved by overcoming the current hardware limitation of the GPU.
\end{abstract}

@@

Transportation systems are part of important infrastructures for our life
and industry.
For example, traffic flow highly influences economy.
According to the report from the Japanese government, economic losses
due to traffic congestion could reach one hundred billion dollars per
year in Japan, which include degradation of transport efficiency, energy
consumption, and environmental poisoning.
Albeit a major issue of transportation systems, the mechanism of traffic
congestion is not well-explained in the literature.
Examples include a ``phantom jam'' -- a traffic jam that naturally
occurs on freeways without any car and/or human accidents.
This phantom jam should be addressed by science, since it is a physical
phenomenon but not one caused by human errors.
There is certainly a challenge of real-time simulation in this line of
work, but even a basic scale of execution times is not understood at all.

@@

Applying a large number of agents to simulation using the above formula,
it is apparent that computational workload increases exponentially.
The above formula considers only one dimension, which restricts
applications of simulation to freeway traffic flows, powder flows,
molecular motors, and so on.
Making it multi-dimensional further increases computational workload,
while allowing more complicated simulations, such as traffic networks,
internet packet flows, evacuation routes, and herd formations of
animals, to use similar optimal velocity models.
Given a scale of million agents in the real world, traffic simulation
must be supported by powerful computer systems.

@@

This paper explores how to accelerate traffic simulation using the
graphics processing unit (GPU), which integrates hundreds of processing
cores on a chip.
Recent GPUs are becoming more and more suitable for general-purpose
data-parallel applications.
The traffic simulation program used in this paper applies
Equation~\eqref{eqn:ov} to a large number of agents.
The resulting workload is highly data-parallel and compute-intensive,
which can be offloaded on to the GPU.
We also identify a bottleneck of traffic simulation when accelerated by
the GPU, and provide an insight into some solutions.
The main objective of this paper is to figure out a computational scale
of traffic simulation.
Timing guarantees are not considered, but our contribution is essential
to achieve real-time traffic simulation in future work.

The rest of this paper is organized as follows.
Section~\ref{sec:assumption} describes the basic assumption and
terminology behind this paper.
Section~\ref{sec:traffic_simulation} explains an overview of
traffic simulation and its literature.
Section~\ref{sec:gpu_implementations} presents our schemes of GPU
implementations for traffic simulation.
Section~\ref{sec:evaluation} evaluates the advantage of using our
schemes, and provides an insight into future work.
Section~\ref{sec:conclusion} concludes this paper.

@@

This paper is particularly focused on CUDA and GPUs, but the notion of
accelerating traffic simulation presented in this paper is applicable to
heterogeneous architectures that integrate CPUs and accelerators.
GPUs are currently the most well-recognized forms of accelerators, but
emerging alternatives include the Intel Many Integrated Core (MIC) and
AMD Fusion architectures.
However, their programming models are almost identical in that a master
thread running on the CPU manages the program flow, while parallelized
threads are offloaded on to accelerators.
CUDA, OpenCL, HMPP, and AMP are all based on this programming model.

@@

So far we have studied a single process of traffic flow.
In reality, however, there are multiple processes of traffic flow.
Real-time feedback from the real road conditions may also be
additionally required to deploy traffic simulation.
All these pieces of simulation must meet the quality-of-service (QoS)
requirement for users.
Imagine how many potential users exist in the real-world transportation
system.
We are now facing a core challenge of next-generation real-time
systems.

Recently real-time systems have been augmented with GPU resource
management techniques~\cite{Elliott_RTS12, Elliott_ECRTS12, Kato2011_3,
Kato2011_2, Kato2011_1, Kato2012, Liu_PACT12}.
Scheduling, synchronization, and memory management of GPU applications
are becoming key challenges to applying the GPU for real-time systems.
A grander vision of GPU-accelerated traffic simulation will need a
tighter integration of real-time systems technology and computational
physics.
We believe that this paper plays a vital role to facilitate this line of
work.

@@

In future work, we extend our scheme to multidimensional OV models.
This extension allows our research to be applied to a broader range of
problems beyond freeways traffic simulation.
We also plan to consider a specific approach to the integration of the
Imprecise Computation model into our scheme in order to facilitate
deployment of traffic simulation in the real world.
This would pose a core challenge of real-time systems.
Another interesting direction of future work is a coordination with
engineering-based simulation such as AutoMatrix~\cite{Mangharam2011}.

\section*{Acknowledgement}

This work was supported by Grant-in-Aid for Scientific Research
24500058.
